Automated Program Verification and Synthesis of Embedded \& Cyber-physical Systems: Advances and Challenges

### 2017-08-08 paper.tex
additional changes to section 3

One promising approach to achieve completeness, in BMC techniques, is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. The main challenge, however, of such an approach relies on computing and strengthening inductive invariants from programs. In particular, loop invariants, which are computed from programs under verification, must be inductive (and not just invariant), in order to check the corresponding VCs, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}.

For instance, consider the C-code fragment shown in Figure~\ref{figure:inductive-invariant}. If we consider the IEEE floating-point standard (IEEE 754)~\cite{4610935,Goldberg91whatevery}, then  the loop invariant $x>0$ holds initially and after each iteration; actually, $x$ tends to infinity after $128$ iterations. This invariant, however, is not inductive, given that $x>0$ before an initial iteration does not ensure that $x>0$ after each iteration; if we initially have $x=0.9$, then $x<0$ after the fourth iteration. As a consequence, even if invariant generation procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that\textit{k}-induction verifiers can automatically prove program correctness. In this specific example, an inductive invariant would be $x>1$, given that if we can check that if $x>1$ before the initial iteration, then also $x>1$ after \textit{k} iterations.
%
\begin{figure}[h]
\centering
\begin{minipage}{0.45\textwidth}
\begin{lstlisting}
#include <assert.h>
int main(void) {
 float x=2;
 while (1) {
   x = ((2*x) - 1);
 }
 return 0;
}
\end{lstlisting}
\end{minipage}
\caption{Motivating example for inductive invariants.}
\label{figure:inductive-invariant}
\end{figure}

@@

Another feasible alternative to prove properties in BMC is to compute the Craig interpolants for inconsistent pairs (or more generally, sets) of formulae~\cite{McMillan03,McMillan05,McMillan06,McMillan07}. This alternative approach exploits the SAT/SMT solvers' ability to produce refutations, {\it i.e.}, proofs that there is no counterexample of depth less than or equal to $k$. This proof does not ensure whether a given property holds in the model, but it contains information about the reachable states of the model.

%\begin{prf}\label{interpolants}
\begin{definition}\label{interpolants}
\textit{Given a pair of formulae $\left(A,B\right)$, and a proof by resolution for $\left(A,B\right)$, an \textit{interpolant} for $\left(A,B\right)$ is a formula $F$ with the following properties~\cite{McMillan03,McMillan05}: }

\begin{itemize}
	\item $A \Rightarrow F$
	\item $F \wedge B$ \textit{is unsatisfiable}
	\item $F$ \textit{refers only to the common variables of $A$ and $B$}
\end{itemize}
\end{definition}
%\end{prf}

As an example, consider $A = \left(x_1 \wedge x_2\right)$ and $B = \left(\neg x_2 \wedge x_3\right)$. Given that $\left(x_1 \wedge x_2\right)$ must imply $F$ (or simply that $\neg x_1 \vee \neg x_2 \vee F$ hold) and $F \wedge \neg x_2 \wedge x_3$ must be unsatisfiable, one possible interpolant for the given pair of formulae $\left(A,B\right)$ is $F = x_2$ since $x_2$ is also common to both $A$ and $B$.

The use of interpolants allows us to define a complete method for finite-state reachability analysis based on SAT and SMT solvers. In order to show how BMC and interpolation can be combined, we refer to Section~\ref{Preliminaries} where we define Eq.~(\ref{bounded-model-checking-biere}) and the terms $I$, $R$, and $\phi$. Now suppose that $Q=I$ and we partition Eq.~(\ref{bounded-model-checking-biere}) so that the set of initial states $I$ and the first instance of the transition relation $R$ are in set $A$, while the remaining instances of $R$ and the property $\phi$ are in set $B$ as shown in Figure~\ref{figure:computing-image-by-interpolation} (note that $k$ is unknown).

@@

Suppose that we use an SMT solver to prove that the $A \wedge B$ is unsatisfiable, {\it i.e.}, we use an SMT solver to conclude that there is no satisfying assignment to $A \wedge B$.\footnote{Note that if at any stage we can satisfy the property $\phi$ within $k$ steps from the initial state, then we have found a counterexample.} The internal steps performed by the SMT solvers for reaching this conclusion can be used to construct a proof of unsatisfiability $\Pi$. From this proof, we can derive an interpolant $F$ for the pair of formulae $\left(A,B\right)$, i.e., $F = interpolant\left(\Pi, A, B\right)$. According to Definition~\ref{interpolants}, $A$ must imply $F$ and since we defined $A$ to be the set of initial states and the first instance of $R$ (i.e., from Figure~\ref{figure:computing-image-by-interpolation}, $A=s_{0} \wedge s_{1}$), it follows that $F$ is \textit{true} in every state reachable from the initial state in one step. In other words, we can say that $F$ is an over-approximation of the forward image of $I$~\cite{McMillan03,McMillan05}. Also according to Definition~\ref{interpolants}, the formula $F \wedge B$ must be unsatisfiable (from Figure~\ref{figure:computing-image-by-interpolation}, $B=s_{2} \wedge s_{3} \wedge \ldots \wedge s_{k}$), which means that there is no state satisfying $F$ that can reach a final state $s_k$. After computing the interpolant $F$, we then check whether $F$ implies $Q$. If $F$ implies $Q$, then no reachable state can satisfy the property $\phi$ and we can thus conclude that the property holds. However, since $F$ is an approximation, we can falsely conclude that the final state is reachable. In this case, we update $Q = F \vee Q$ and $A = F \wedge R_{0}$, increase the value of $k+1$ and check whether $A \wedge B$ is unsatisfiable. If $A \wedge B$ is satisfiable, we have found a valid counter-example (i.e., a path from the initial state to the final state). Otherwise, we compute the interpolant $F = interpolant\left(\Pi, A, B\right)$ again and check whether $F$ implies $Q$. We stop this procedure when we have found a valid counterexample or have proved that the final state is not reachable (i.e., the property holds). The details of the algorithm and further information about the use of interpolants in model checking can be found in~\cite{McMillan03,McMillan05,McMillan06,McMillan07}.

@@

In summary, DSVerifier is a useful test tool, which takes into account different representations ({\it e.g.}, transfer-function and state-space), realization forms ({\it e.g.}, direct, delta, and transposed forms), and other implementation restrictions, in order to explore the design space. Finally, if a system's requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup, which may even be incorporated into a given system development procedure, by providing both system verification and model refinement.

@@

One may also notice that such a paradigm, {\it i.e.}, verification and synthesis based on an expected system behavior, is not restricted to digital filters and controllers or DMA races, but it can also be applied to possibly any real system, as long as the desired behavior can be expressed as properties in BMC frameworks. For instance, regarding self-driving cars, an important property could be the detection of pedestrians, animals, bicycles, or any obstacle present on urban and rural roads. Indeed, for instance, bicycles are considered one of the most difficult problems, due to the myriad of possible shape and colors \cite{selfcar}. In addition, the mentioned problem is closely related to machine learning, from which refinement strategies can be devised and incorporated to BMC approaches \cite{BMCml}.

>>>>

One promising approach to achieve completeness, in BMC techniques, is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}; however, the main challenge regarding such an approach relies on computing and strengthening inductive invariants from programs. In particular, loop invariants, which are computed from programs under verification, must be inductive (and not just invariant), in order to check the corresponding VCs, {\it i.e.}, induction cannot determine invariance of a non-inductive assertion~\cite{Bradley07}.

For instance, consider the C-code fragment shown in Figure~\ref{figure:inductive-invariant}. If we consider the IEEE floating-point standard (IEEE 754)~\cite{4610935,Goldberg91whatevery}, then  the loop invariant $x>0$ holds initially and after each iteration; actually, $x$ tends to infinity after $128$ iterations. Nonetheless, this invariant is not inductive, given that $x>0$ before an initial iteration does not ensure that $x>0$ after each iteration; if we initially have $x=0.9$, then $x<0$ after the fourth iteration. As a consequence, even if invariant generation procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that \textit{k}-induction verifiers can automatically prove program correctness. In this specific example, an inductive invariant would be $x>1$, given that if we can check that if $x>1$ before the initial iteration, then also $x>1$ after \textit{k} iterations.
%
\begin{figure}[h]
\centering
\begin{minipage}{0.45\textwidth}
\begin{lstlisting}
#include <assert.h>
int main(void) {
 float x=2;
 while (1) {
   x = ((2*x) - 1);
 }
 return 0;
}
\end{lstlisting}
\end{minipage}
\caption{Motivating example for inductive invariants.}
\label{figure:inductive-invariant}
\end{figure}

@@

Another feasible alternative to prove properties in BMC is to compute Craig interpolants for inconsistent pairs (or more generally, sets) of formulae~\cite{McMillan03,McMillan05,McMillan06,McMillan07}. This alternative approach exploits the SAT/SMT solvers' ability to produce refutations, {\it i.e.}, proofs regarding the nonexistence of counterexamples of depth less than or equal to $k$. Those proofs do not ensure whether a given property holds in the model, but they contain information about reachable states of a model.

%\begin{prf}\label{interpolants}
\begin{definition}\label{interpolants}
\textit{Given a pair of formulae $\left(A,B\right)$, such that $A \wedge B$ is inconsistent, and a proof by resolution for $\left(A,B\right)$, an \textit{interpolant} for $\left(A,B\right)$ is a formula $F$ with the following properties~\cite{McMillan03,McMillan05}: }

\begin{itemize}
	\item $A \Rightarrow F$;
	\item $F \wedge B$ \textit{is unsatisfiable};
	\item $F$ \textit{expressed only over the common variables (nonlogical symbols) of $A$ and $B$}.
\end{itemize}
\end{definition}
%\end{prf}

As an example, consider $A = \left(x_1 \wedge x_2\right)$ and $B = \left(\neg x_2 \wedge x_3\right)$. Given that $\left(x_1 \wedge x_2\right)$ must imply $F$ (or simply that $\neg x_1 \vee \neg x_2 \vee F$ hold) and $F \wedge \neg x_2 \wedge x_3$ must be unsatisfiable, one possible interpolant for the given pair of formulae $\left(A,B\right)$ is $F = x_2$, since $x_2$ is a common part of both $A$ and $B$.

The use of interpolants allows us to define a complete method for finite-state reachability analysis based on SAT and SMT solvers. In order to show how BMC and interpolation can be combined, we refer to Section~\ref{Preliminaries} where we define Eq.~(\ref{bounded-model-checking-biere}) and the terms $I$, $R$, and $\phi$. Now, suppose that $Q=I$ and Eq.~(\ref{bounded-model-checking-biere}) is partitioned, so that the set of initial states $I$ and the first instance of the transition relation $R$ are in set $A$, while the remaining instances of $R$ and the property $\phi$ are in set $B$, as shown in Figure~\ref{figure:computing-image-by-interpolation} (note that $k$ is unknown).

@@

Suppose that we use an SMT solver to prove that the $A \wedge B$ is unsatisfiable, {\it i.e.}, we use an SMT solver to conclude that there is no satisfying assignment to $A \wedge B$.\footnote{Note that if at any stage we can satisfy the property $\phi$ within $k$ steps from the initial state, then we have found a counterexample.} The internal steps performed by the SMT solvers for reaching this conclusion can be used to construct a proof of unsatisfiability $\Pi$. From this proof, we can derive an interpolant $F$ for the pair of formulae $\left(A,B\right)$, i.e., $F = interpolant\left(\Pi, A, B\right)$. According to Definition~\ref{interpolants}, $A$ must imply $F$ and since we defined $A$ to be the set of initial states and the first instance of $R$ (i.e., from Figure~\ref{figure:computing-image-by-interpolation}, $A=s_{0} \wedge s_{1}$), it follows that $F$ is \textit{true} in every state reachable from the initial state in one step. In other words, we can say that $F$ is an over-approximation of the forward image of $I$~\cite{McMillan03,McMillan05}. Also according to Definition~\ref{interpolants}, the formula $F \wedge B$ must be unsatisfiable (from Figure~\ref{figure:computing-image-by-interpolation}, $B=s_{2} \wedge s_{3} \wedge \ldots \wedge s_{k}$), which means that there is no state satisfying $F$ that can reach a final state $s_k$. After computing $F$, we then check whether $F$ implies $Q$. If $F$ implies $Q$, then no reachable state can satisfy the property $\phi$ and we can thus conclude that the property holds; however, since $F$ is an approximation, we can falsely conclude that the final state is reachable. In this case, we update $Q = F \vee Q$ and $A = F \wedge R_{0}$, increase the value of $k+1$ and check whether $A \wedge B$ is unsatisfiable. If $A \wedge B$ is satisfiable, we have found a valid counter-example (i.e., a path from the initial state to the final state); otherwise, we compute the interpolant $F = interpolant\left(\Pi, A, B\right)$ again and check whether $F$ implies $Q$. This procedure is stopped when we have found a valid counterexample or have proved that the final state is not reachable (i.e., the property holds). The details of the algorithm and further information about the use of interpolants in model checking can be found in~\cite{McMillan03,McMillan05,McMillan06,McMillan07}.

@@

In summary, DSVerifier is a useful test tool, which takes into account different representations ({\it e.g.}, transfer-function and state-space), realization forms ({\it e.g.}, direct, delta, and transposed forms), and other implementation restrictions, in order to explore the design space. Finally, if a system's requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup, which can even be incorporated into a given system development procedure, by providing both system verification and model refinement.

@@

One may also notice that such a paradigm, {\it i.e.}, verification and synthesis based on an expected system behavior, is not restricted to digital filters and controllers or DMA races, but it can also be applied to possibly any real system, as long as the desired behavior can be expressed as properties in BMC frameworks. For instance, regarding self-driving cars, an important property could be the detection of pedestrians, animals, bicycles, or any obstacle present on urban and rural roads. Indeed, bicycles are considered one of the most difficult problems, due to the myriad of possible shape and colors \cite{selfcar}. In addition, the mentioned problem is closely related to machine learning, from which refinement strategies can be devised and incorporated to BMC approaches \cite{BMCml}.
